### 2019-08-13 cover-letter-jmlr.tex
Polishing by CJ

\documentclass[a4paper]{article}

\usepackage{a4wide} 
\usepackage{subfigure} 
\usepackage{psfrag}
\usepackage{amsmath}
\usepackage{url}
\usepackage{natbib}
\usepackage{color}
\usepackage{ulem}
\normalem

\usepackage{parskip}

\usepackage{sectsty}
\sectionfont{\large}
\subsectionfont{\normalsize}

\setlength{\textwidth}{6.1in}

\bibliographystyle{unsrt}

\begin{document}

\begin{center}
{\Large \bf metric-learn: Metric Learning Algorithms in Python\\\vspace{.5cm}Cover Letter for JMLR MLOSS Submission}
\vspace{1cm}
\end{center}

Dear JMLR editors,\\

Our submission ``metric-learn: Metric Learning Algorithms in Python'' is intended for the machine learning open source software section. We give here the required information about the software:
\begin{itemize}
    \item Name: metric-learn
    \item License: MIT License
    \item Project URL: \url{https://github.com/scikit-learn-contrib/metric-learn}
    \item Version to be reviewed: v0.5.0\\
\end{itemize}

We look forward to hearing your feedback on our submission.\\

Best regards,\\

William de Vazelhes, CJ Carey, Yuan Tang, Nathalie Vauquier and Aur\'elien Bellet

\end{document}