### 2020-07-01 cover-letter-jmlr-revision.tex
Updates from Overleaf

\documentclass[a4paper]{article}

\usepackage{a4wide} 
\usepackage{subfigure} 
\usepackage{psfrag}
\usepackage{amsmath}
\usepackage{url}
\usepackage{natbib}
\usepackage{color}
\usepackage{ulem}
\normalem

\usepackage{parskip}

\usepackage{sectsty}
\sectionfont{\large}
\subsectionfont{\normalsize}

\setlength{\textwidth}{6.1in}

\bibliographystyle{unsrt}

\begin{document}

\begin{center}
{\Large \bf metric-learn: Metric Learning Algorithms in Python\\\vspace{.5cm}Cover Letter for JMLR MLOSS Submission}
\vspace{1cm}
\end{center}

Dear JMLR editors,\\

We have the honor to submit the revised version of the manuscript, entitled "metric-learn: Metric Learning Algorithms in Python" for publication in Journal of Machine Learning Research. All the comments made by the referee have been addressed.

Best regards,\\

William de Vazelhes, CJ Carey, Yuan Tang, Nathalie Vauquier and Aur\'elien Bellet

\end{document}