### 2020-07-01 header-PBP.tex

%%%%% DOCUMENT SETUP
\usepackage[top = 1.5cm, bottom = 1.5cm, right = 1.5cm, left = 1.5cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[titles]{tocloft}
\usepackage[explicit]{titlesec}
\usepackage{fancybox}
\usepackage[bottom]{footmisc}
\usepackage{fancyhdr}
\usepackage{etex}
\usepackage{lastpage}
\usepackage{setspace}

%%%%% FONTS
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[sc]{mathpazo}
\usepackage{dsfont}

%%%%%% SYMBOLS
%\usepackage{latexsym, pifont}

%%%%% MATHS
\usepackage{amssymb, amsfonts, amsmath, euler, amscd, amsthm, array}

%%%%%% TABLES
\usepackage{booktabs, tabularx, float, subfig, multicol, multirow, caption, wrapfig}
\setlength{\aboverulesep}{0pt}
\setlength{\belowrulesep}{0pt}
\captionsetup{format = plain, labelfont = normalfont, textfont = {it, small}, labelsep = endash, justification = centerlast}
\usepackage{color, colortbl}
\usepackage[table]{xcolor}

%%%%%% FIGURES
\usepackage{graphicx, epsfig, tikz, subfig}

%%%%%% HYPERREF
\RequirePackage[colorlinks, citecolor = autumnred, urlcolor = autumnred, linkcolor = darkautumnred]{hyperref}

%%%%% LISTS
%\usepackage{enumerate}
\usepackage{enumitem}

%%%%% MISC
\usepackage{pdfpages}
\usepackage{textcomp}
\usepackage{sidecap}
\usepackage{ifthen}

%%%%% TITLES FORMAT
\renewcommand{\thesection}{\Roman{section}}
\renewcommand{\thesubsection}{\arabic{subsection}}
\titleformat{\section}[hang]{\centering\scshape\LARGE\color{bordeau!50!darkautumnred}}{\huge\thesection\,-\,}{0pt}{#1\\[-1em]\textcolor{bordeau!50!darkautumnred}{\hrule height 1.5pt}}
\titleformat{\subsection}[hang]{\scshape\large}{{\bfseries\color{bordeau!50!darkautumnred}\thesubsection \,-\,}}{5pt}{#1}{}
%\titleformat{\paragraph}[hang]{\scshape}{}{5pt}{#1}{}

%%%%%% TITLES FORMAT
%\renewcommand{\thesection}{\huge \roman{section}\,-\,}
%\renewcommand{\thesubsection}{\arabic{subsection}\,-\,}
%\titleformat{\section}[hang]{\scshape\LARGE\color{bordeau}}{\thesection}{0pt}{#1 \textcolor{bordeau}{\titlerule[1.25pt]}}
%\titleformat{\subsection}[hang]{\bfseries\large}{\thesubsection}{5pt}{ #1}

%%%%% COLORS
\definecolor{bordeau}{rgb}{0.46 , 0.06 , 0.06}
\definecolor{gris}{rgb}{0.60 , 0.60 , 0.60}
\definecolor{grisclair}{RGB}{234,234,234}
\definecolor{autumnred}{RGB}{169,56,38}
\colorlet{darkautumnred}{autumnred!120}

%%%%% NEW COMMANDS : SHORTCUTS

%% LATIN ABBREVIATIONS
\newcommand{\ie}{\textit{i.e.}\hspace{4pt}}
\newcommand{\eg}{\textit{e.g.}\hspace{4pt}}
\newcommand{\cf}{\textit{cf.}\hspace{4pt}}

\newcommand{\red}[1]{\textcolor{red}{#1}}

%% MATH SYMBOLS
\newcommand{\opset}[1]{ \overset{\circ}{#1} } % Interior
\newcommand{\wideopset}[1]{ \overset{\circ}{\overgroup{#1}} } % Interior (wide)
\newcommand{\clset}[1]{ \overline{#1} } % Closure
\newcommand{\ginv}[1]{ #1^{\leftarrow} } % Generalized inverse
\newcommand{\Prob}[2][]{ \mathbb{P}_{#1}\left( #2 \right)} % Probability
\newcommand{\Probhat}[2][]{ \widehat{\mathbb{P}}_{#1}\left( #2 \right)} % Probability estimator
\newcommand{\Esp}[2][]{ \mathbb{E}_{#1}\left( #2 \right) } % Expectation
\newcommand{\Esphat}[2][]{ \widehat{\mathbb{E}}_{#1}\left( #2 \right) } % Expectation estimator
\newcommand{\Var}[2][]{ \mathbb{V}_{#1}\left( #2 \right) } % Variance
\newcommand{\Varhat}[2][]{ \widehat{\mathbb{V}}_{#1}\left( #2 \right) } % Variance estimator
\newcommand{\ascv}[1][n]{ \underset{ {#1} \rightarrow \infty }{\longrightarrow} } % Asymptotic convergence
\newcommand{\ascvvg}[1][n]{ \overset{v}{\underset{ {#1} \rightarrow \infty }{\longrightarrow} }} % Vague asymptotic convergence
\newcommand{\ascvlaw}[1][n]{ \overset{\mathcal{D}}{\underset{ {#1} \rightarrow \infty }{\longrightarrow}} } % Asymptotic convergence in distribution
\newcommand{\ascvprob}[1][n]{ \overset{\mathbb{P}}{\underset{ {#1} \rightarrow \infty }{\longrightarrow}} } % Asymptotic convergence in probability
\newcommand{\iid}{\overset{iid}{\leadsto}} % iid distributed
\newcommand{\vlim}[1][n]{\underset{#1 \rightarrow \infty}{v\hspace{-1.5pt}\lim} \,\, } % Vague limit
\newcommand{\indic}[1]{\mathbb{I}\left\{ #1 \right\}} % Indicator function
\newcommand{\COND}{\bigg\vert} % Conditionning (large separator)

%% MATH FONTS
\def\P{\mathcal P}
\def\S{\mathcal S}
\def\RR{\mathbb R}
\newcommand{\F}{{\cal F}}
\newcommand{\G}{{\cal G}}
\newcommand{\X}{{\cal X}}
\newcommand{\D}{{\cal D}}
\newcommand{\cR}{{\cal R}}
\def\C{{\cal C}}
\def\B{{\cal B}}
\def\A{{\cal A}}
\def\N{{\cal N}}
\def\F{{\cal F}}
\def\S{{\cal S}}
\def\L{{\cal L}}
\def\K{{\cal K}}

%%% DOCUMENT INFORMATIONS
% AUTHORS
\newcommand{\theauthor}{}	
\newcommand{\theshortauthor}{}	% Short version optional
\renewcommand*{\author}[2][]{\renewcommand{\theshortauthor}{#1}\renewcommand{\theauthor}{#2}}
% JOURNAL
\newcommand{\thejournal}{}	
\newcommand{\theshortjournal}{}	% Short version optional
\newcommand{\journal}[2][]{\renewcommand{\theshortjournal}{#1}\renewcommand{\thejournal}{#2}}
% SUBMISSION NUMBER
\newcommand{\thesubnb}{}
\newcommand{\subnb}[1]{\renewcommand{\thesubnb}{#1}}
% DATE
\newcommand{\thedate}{}
\renewcommand{\date}[1]{\renewcommand{\thedate}{#1}}
% PAPER TITLE
\newcommand{\thepaper}{}
\newcommand{\theshortpaper}{}
\newcommand{\paper}[2][]{\renewcommand{\theshortpaper}{#1}\renewcommand{\thepaper}{#2}}

%%% LENGHTS
\setlength{\parskip}{4pt}				%Saut de paragraphe
\setlength{\headheight}{15.2pt}			%Hauteur de l'entête

%%% HEADER & FOOTER
\pagestyle{fancy}
\fancyhead{}						
\fancyhead{\begin{tikzpicture}
[remember picture, overlay] \node[anchor = north west, fill = gray!50, minimum width = \paperwidth, text width = .95\paperwidth, text ragged, inner sep = 5pt] at (current page.north west) {\itshape \ifthenelse{\equal{\theshortauthor}{}}{\theauthor}{\theshortauthor} \hfill \ifthenelse{\equal{\theshortpaper}{}}{\thepaper}{\theshortpaper}}; \end{tikzpicture}}
\fancyfoot{}						
\fancyfoot{\begin{tikzpicture}
[remember picture, overlay] \node[anchor = south west, fill = gray!50, minimum width = \paperwidth, text width = .95\paperwidth, text ragged, inner sep = 5pt] at (current page.south west) {{\itshape \thedate} \hfill \thepage /\pageref{LastPage}}; \end{tikzpicture}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

%%% TITLE LAYOUT
\renewcommand{\maketitle}{
%\shorthandoff{:}\shorthandoff{!}
\begin{tikzpicture}
	\node[minimum width = \paperwidth, minimum height = 0.4\paperheight] at (current page.north west) {};
\end{tikzpicture}
\begin{tikzpicture}[remember picture, overlay]
%
%	\node[fill = gray!50, minimum width = \paperwidth, text width = 0.95\paperwidth, inner xsep = 1em, inner ysep = 1em, anchor = north west] (head) at (current page.north west) {\large \scshape \theschool \hfill \thelevel};
%
	\node[fill = gray, minimum width = \paperwidth, inner ysep = 3em, anchor = north west] (title) at ([yshift = -7em]current page.north west) {\begin{minipage}{.95\paperwidth} \begin{spacing}{2}
\centering {\huge \scshape Point-by-point reply to the referee report on \thejournal \ submission No. \thesubnb}
	\end{spacing}\vspace*{-1em}
	\end{minipage}};
%	
%	\node[anchor = base, minimum width = \paperwidth] (course) at  ($(head.south)!0.5!(title.north)$) {\begin{minipage}{\paperwidth} \centering {\scshape \Huge \thecourse} \ifthenelse{\equal{\thesubtitle}{}}{}{{\\\vspace*{2em} \Large \slshape Cours de \theprof}} \end{minipage}};
%	
%	\node[fill = gray!50, minimum width = \paperwidth, text width = .95\paperwidth, inner xsep = 1em, inner ysep = 1em, anchor = south west] (foot) at (current page.south west) {\large  \scshape \theauthor \hfill \thedate};
%	
	\node[anchor = north west, minimum width = \paperwidth] (auth) at ([xshift = 2em, yshift = -3em]title.south west) {\begin{minipage}{\paperwidth} {\scshape\color{bordeau}\Large Authors:} \hspace{.5em} {\large\textbf{\theauthor}} \end{minipage}};
	\node[anchor = north west, minimum width = \paperwidth] (pap) at ([yshift = -1em]auth.south west) {\begin{minipage}{\paperwidth} {\scshape\color{bordeau}\Large Title:} \hspace{.5em} {\large\textbf{\thepaper}} \end{minipage}};
%
\end{tikzpicture}
%\shorthandon{:}\shorthandon{!}
}